---
title: "Producing caterpillar plots for univariate models"
author: "Forest Hayes"
output:
  prettydoc::html_pretty:
    theme: cayman

---


####The goal here is to produce caterpillar plots to compare univariate model results from two different sampling methods for three data sets. Since we have nine different covariates that means we have to run and plot a whopping 54 models.  Here I make  use of lists and apply functions in order to facilitate the process - hopefully I have also provided enough annotations to make it easy to follow.



# 1. Running models

Let's assume we already have our used and available points and the associated H.S.I. index values in a list of two for MCP and KDE.


```{r, include = F}
require(magrittr)                                     # load pipe opperator
require(ggplot2)                                      # load ggplot
load("wolfdata.rdat")
```


```{r, results = "asis"}
wolf_data %>%                                         # look at the data
purrr::map(~head(., n = 2) %>% knitr::kable(format = "markdown"))
```

<br>

Since we need to run 54 different models, we can greatly facilitate the process by using lists and apply functions.

First, lets make a function to run models and return a tidy dataframe as a result.

```{r}

# FUNCTION to produce tidy glm results
tidy_glm <- function(data = data, x = x, y = y){
  glm(                                                # GLM function
    formula = as.formula(paste(y, "~", x, sep = " ")),  # specify glm formula
    family = binomial(logit),                           # logit function
    data = data                                         # specify data
  ) %>% 
  broom::tidy(exponentiate = T) %>%                   # tidy using broom
  dplyr::slice(2)                                     # return only coefficients
}

```

Now we can pass our data, and specify x and y arguments for the function. Note that we use <tt>broom::tidy()</tt> to produce a data frame of our covariate estimates and can use the argument <tt>exponentiate = T</tt> to convert the estimates from the log to probability scale. Cool!

Lets test it out:

```{r}

tidy_glm(data = wolf_data$KDE, x = "Deer", y = "used") %>% 
knitr::kable()

```


This is pretty good... but we really want to be able to give the function a whole data frame an get a list of results back so lets make another function that builds off the first one.

```{r}
# FUNCTION to apply tidy_glm() to each col of 'data' using a list of 'x' cols
tidy_glms <- function(data = data, x = x, y = "used", model = "model"){
  purrr::map(                                         # lapply
    .f = tidy_glm,                                    # tidy glm function
    .x = x,                                           # for each value of x
    data = data,                                      
    y = y                                             # for each (or 1) val of y 
  ) %>% 
  dplyr::bind_rows() %>%                              # bind rows of results
  dplyr::mutate(model = model)                        # add column for 'model'
}

# Note that this function also has an argument to add a common column that 
# specifies the model used to produce the data (in this case MCP or KDE).

```

<br>

Let's get a list of the column names we want to analyze to use in our model.
```{r}
var_names <-
  wolf_data$KDE %>%                                   # pass data
  dplyr::select(-Pack, -used) %>%                     # remove grouping cols
  colnames() %>%                                      # get colnames
  as.list()                                           # cnvrt to list
```


And subset the data for each pack.
```{r}
red_deer_list <-  
  wolf_data %>% 
  purrr::map(~dplyr::filter(., Pack == "Red Deer")) 

bow_valley_list <-
  wolf_data %>% 
  purrr::map(~dplyr::filter(., Pack == "Bow Valley")) 
```


And run the full set of models for each pack group.
```{r, results = "hide"}

# red deer pack model results
red_deer_res <- 
  purrr::map2(
    .x = red_deer_list,
    .y = list("KDE", "MCP"),
    .f = function(df, model)
      tidy_glms(data = df, x = var_names, y = "used", model = model)
  ) %>% 
  dplyr::bind_rows() 

# bow valley pack model results
bow_valley_res <- 
  purrr::map2(
    .x = bow_valley_list,
    .y = list("KDE", "MCP"),
    .f = function(df, model)
      tidy_glms(data = df, x = var_names, y = "used", model = model)
  ) %>% 
  dplyr::bind_rows() 

# combined pack model results
combined_pack_res <- 
  purrr::map2(
    .x = wolf_data,
    .y = list("KDE", "MCP"),
    .f = function(df, model)
      tidy_glms(data = df, x = var_names, y = "used", model = model)
  ) %>% 
  dplyr::bind_rows() 
```


# 2. Creating plots

Now that we have the model results we can actually make some plots!  The <tt>dotwhisker</tt> package is a wrapper for <tt>ggplot</tt> that uses tidy model results to produce caterpillar plots. Lets look at one of our model sets to start.

```{r}
red_deer_res %>% 
dotwhisker::dw_plot()
```

Pretty good! Let's tweak the aesthetics of the plot and wrap it into a function.
```{r, results= "hide"}
# FUNCTION to produce dot whisker plots
boxwhiskerplot <- function(df, title = "Comparison of covariate estimates"){
  df %>% 
  dotwhisker::dw_plot(                                # create dot whisker plot
    dodge_size = 0.7,
    dot_args = list(shape = rep(c(15, 16), 9), size = 2.4),  # change dot shapes
    whisker_args = list(size = 1)                            # whisker size
  ) +
  guides(                                             # change legend symbols
    color = guide_legend(                          
      override.aes = list(size = 5, shape = c(16, 15)))
  ) +
  theme_bw() +                                        # set theme to bw
  ggtitle(title) +                                    # add title
  xlab("Effect size") +
  theme(                                              # modify theme
    plot.title = element_text(                          
      size = 12,                                        # title font size
      face = "bold"                                     # bold text
    ),
    panel.grid.major = element_blank(),                 # remove major grids
    panel.grid.minor = element_blank()                  # remove minor grids
  ) +
  scale_x_continuous(                                 # change xlim 
    limits=c(0, 4.5),                                   # set limits
    expand = c(0.01, 0.01)                              # padding past limits
  ) +
  geom_vline(                                         # add vertical line
    xintercept = 1,                                     # at x= 1
    lty = "dotted"                                      # line type
  ) +
  scale_color_manual(                                 # change color + label
    labels = c("KDE", "MCP"),                           # labels
    values = c("#e41a1c", "#377eb8")                    # colors
  ) +
  labs(color = "Method")                              # change legend title
}
```



Now apply that function to each of the model result tables to produce  our plots.
```{r}
plot_list <-
  purrr::map2(                                          
    .x = list(                                        # list of pack data
      combined_pack_res,
      bow_valley_res,
      red_deer_res),
    .y = list(                                        # list of titles
      "Comparison of covariate estimates for both wolf packs combined",
      "Bow Valley wolf pack comparison of covariate estimates",
      "Red Deer wolf pack comparison of covariate estimates"
    ),
  .f = function(df, title)
    boxwhiskerplot(df, title)                         # plot function
  )
```


Finally, let's put them into a single three part figure.

```{r}
out <-
  ggpubr::ggarrange(                                    # arrange into grid
    plotlist = plot_list,
    common.legend = T,                                  # use a common legend
    legend = "bottom",                                  # legend position
    labels = "AUTO",                                    # label with LETTERS
    ncol = 1,
    nrow = 3
  )
```

## Final product:

<br>

```{r, echo = F, fig.height = 8, fig.width = 8, fig.cap= "Figure 1: Comparison of logit link binomial regression coefficient estimates using kernel density home range estimates (KDE) and minimum convex polygon (MCP) estimates. For each species, the x-axis represents the an increase of H.S.I. value.  For DistHighHuman, DistHuman, and Elevation, a change on one on the x-axis represents an increase of 100 m. For each covariate the mean is plotted with 95% confidence intervals. A) shows model results for the combination of both packs. In this model, confidence intervals overlapped for all covariates.  B) shows the estimates for the Bow Valley Pack. Here, the KDE estimate was substantially higher for Wolf, Deer, and Elk than the MCP estimates. C) shows model results for the Red Deer pack. Estimates were more similar for this pack than the Bow Valley pack, however the KDE model still estimated larger effects for changes in Wolf, Deer, and Elk, H.S.I."}

out
```










